Inverse design workflow of implicit surface-based cellular materials via conditional generative model and active learning

### 2025-02-01 manuscript.tex
update on mac 2025-01-27
update on mac 2025-0201

\title{Inverse design workflow of implicit surface-based cellular materials via conditional generative model and active Learning}

\author[1]{Jiaxuan Ma}
\author[3]{Bin Cao}
\author[1]{Yuan Tian, \corref{cor1}}
\ead{tian0321@shu.edu.cn}
\author[1,2]{Sheng Sun, \corref{cor1}}
\ead{mgissh@t.shu.edu.cn}

@@

\begin{abstract}
Implicit surface-based cellular materials have gained significant attention in bone scaffold applications due to their structural resemblance to natural bone. These materials exhibit excellent mechanical properties and bioadaptability, providing an ideal environment for bone tissue regeneration. However, the rational design of such materials often relies heavily on expert knowledge and extensive trial-and-error experiments. In the work, we propose a data-efficient, two-stage bi-objective optimization framework within a machine learning paradigm, which accounts for the disparate computational costs associated with different objectives. Specifically, we apply this methodology to the inverse design of orthopedic implants, integrating a generative model, active learning, and finite element analysis to achieve cellular materials with both biocompatible elastic modulus and enhanced yield strength. Our approach demonstrates significant improvements in overall performance compared to the initial dataset, while reducing design costs by approximately sixfold relative to conventional one-step bi-objective machine learning optimization methods. This framework establishes an efficient paradigm for the rapid and intelligent design of cellular materials with tailored mechanical properties, particularly in scenarios where objectives involve unbalanced computational costs.
\end{abstract}

@@

%%Research highlights
\begin{highlights}
\item A TPMS-GAN with a navigator was developed to generate TPMS unit cells under desired mechanical and geometric conditions.
\item An Inverse TPMS-GAN was introduced to design TPMS unit cells directly from geometric features in an explainable way, bypassingneed to interpolate in latent space or modify implicit function coefficients.
\item A constraint-aware active learning method was proposed to address bi-objective optimization problems with unbalanced objective function costs.
\end{highlights}

@@

In the work, we propose a two-step strategy that integrates generative model with active learning algorithms to address bi-objective design problem of TPMS unit cells associated with imbalanced costs of properties. As demonstrated in Fig \ref{fig:1}, our approach consists of three main parts: (1) generate structures with a desired elastic modulus. In this step, a TPMS-GAN model is trained on a dataset containing TPMS unit cells and their corresponding elastic modulus. Then the generator of the TPMS-GAN produces structures with target elastic modulus, forming a virtual structure space (VSS). (2) inversed TPMS-GAN (ITPMS-GAN) model was developed to explore the latent space of TPMS-GAN, allowing for direct geometric manipulation of TPMS structures instead of modifying implicit function parameters. (3) an elastic modulus constraint-aware active learning approach for yield strength optimization. Initially, the yield strength of a select number of TPMS structures within VSS are evaluated to establish the initial dataset via FEM. Subsequently, elastic modulus constraint-aware active learning based on Bayesian optimization (BO) is employed to optimize the yield strength of TPMS unit cells. BO method has been proven significant efficiency in materials design with small data \citep{Ma2024b,Cao2024a,Tian2024a}.

\begin{figure}
    \centering
    \includegraphics[width=1\linewidth]{figures/1.pdf}
    \caption{\textbf{An overview of the proposed workflow for TPMS unit cells design.} \textbf{a} The generator in TPMS-GAN model produces virtual design space ($\Omega_E$) with desired elastic modulus. \textbf{b} The ITPMS-GAN model was developed to edit the TPMS structure through geometry features to rich diversity of candidates in $\Omega_E$. \textbf{c} The machine learning (ML) algorithm interactively queries the FEM to discovery new TPMS designs with high yield strength and target elastic modulus through constrained active learning in $\Omega_E$.}
    \label{fig:1}
\end{figure}

@@

In the work, a voxel-based asymptotic homogenization method \citep{Dong2019} is employed to calculate the effective elastic tensor of TPMS unit cells. The asymptotic homogenization method posits that physical quantities such as stress and strain depend on both the macroscale ($X$) and the microscale ($x/\varepsilon$), where $\varepsilon$ is the scale factor. These quantities are homogenized at the macroscale and exhibit periodicity at the microscale. The microscopic strain tensor can be asymptotically expanded, and by neglecting higher-order terms, it can be expressed as:
\begin{equation}
    \varepsilon = \bar{\varepsilon} + \varepsilon^{*}
\label{eq:2}
\end{equation}
Here, $\bar{\varepsilon}$ represents the mean or macroscopic strain within the unit cell, while $\varepsilon^{*}$ denotes the perturbation strain at the periodic microscale. The perturbation strain is derived from the mechanical balance equation:
\begin{equation}
\int_{\Omega}\hat{\varepsilon}^T:\mathbb{C}:\varepsilon^{*}d\Omega = \int_{\Omega}\hat{\varepsilon}^T:\mathbb{C}:\bar{\varepsilon}
\label{eq:3}
\end{equation}
In this equation, $\hat{\varepsilon}$ is the virtual strain, $\mathbb{C}$ is the elastic tensor of the constituent materials, and $\Omega$ represents the unit cell. This equation is numerically discretized and evaluated using the FEM, which employs eight-node hexahedrons to voxelize each unit cell into $40 \times 40 \times 40$ cubes (more details are provided in Supplement S1). Once $\varepsilon^{*}$ is obtained, the effective elastic tensor $\mathbb{C}^H$ can be determined using the following equation:
\begin{equation}
\mathbb{C}^H:\bar{\varepsilon} = \frac{1}{|\Omega|}\int_{\Omega}\mathbb{C}:(\bar{\varepsilon}+\varepsilon^{*})d\Omega
    \label{eq:4}
\end{equation}
where $|\Omega|$ denotes the volume of the unit cell.

@@

In this matrix, $E_i$ represents the elastic modulus along the $i^{\text{th}}$ direction, where $i \in \{x, y, z\}$. In the work, for cubic symmetric unit cells, we have effective elastic module $E^H=E_x=E_y=E_z$. Additionally, elastic moduli in other directions can be obtained from $\mathbb{S}^H$ using the following equation:
\begin{equation}
\frac{1}{E_{ijk}} = \mathbb{S}^H_{11} - 2 \left(\mathbb{S}^H_{11} - \mathbb{S}^H_{12} - \frac{1}{2} \mathbb{S}^H_{44} \right) \times \left( \ell_{i1}^2 \ell_{j2}^2 + \ell_{j2}^2 \ell_{k3}^2 + \ell_{i1}^2 \ell_{k3}^2 \right)
    \label{eq:6}
\end{equation}
Here, $E_{ijk}$ is the elastic moduli along the $[i,j,k]$ direction, and $\ell$ is the cosine of the angle between the orientation $[i,j,k]$ and the coordinate axes $[x, y, z]$. This approach allows for the evaluation of elastic modulus in any orientation of the unit cell, facilitating the construction of a three-dimensional elastic modulus surface.

@@

In the work, we employ a conditional Wasserstein GAN comprising four components: a generator network ($G_\phi$), a discriminator network ($D_\theta$),  a navigator network ($R_\omega$) and a condition module $\boldsymbol{C}$ that consists of mechanical properties and geometry characterization of the TPMS unit cell. Each condition can be turned on/off during the training process to meet various design requirements (see Fig.\ref{fig:3} ). The generator $G_\phi$ (with set of trainable parameters $\phi$) map the Gaussian noise ($z$) and conditions to TPMS's parameter space $\boldsymbol{p}$ in $\mathbb{R}^4$, and achieve the fake $\boldsymbol{p}'$. While the discriminator $D_\theta$ (with set of trainable parameters $\theta$) evaluate whether a given set of structure parameters are "real" or not. It takes $\boldsymbol{p},\boldsymbol{p}'$ and  $\boldsymbol{C}$ as input and its output $D_\theta(\boldsymbol{p}'|\boldsymbol{C})$ is a function value representing the degree of realism of $\boldsymbol{p}'$. During the training process, the goal of the generator $G_\phi$ is to produce structure parameters $\boldsymbol{p}'$ of TPMS that can deceive the discriminator $D_\theta$. Simultaneously, the objective of $D_\theta$ is to distinguish structures $\boldsymbol{p}'$ generated by $G_\phi(\boldsymbol{z}|\boldsymbol{C})$ and real structure $\boldsymbol{p}$. $G_\phi$ and $D_\theta$ play the following two-player minimax game with loss function:
\begin{equation}
\mathcal{L}_{(G_\phi, D_\theta)}= \min_G \max_{D} \mathbb{E}_{\boldsymbol{p} \sim \mathbb{P}_r} \left[ D_\theta(\boldsymbol{p}|\boldsymbol{C}) \right] - \mathbb{E}_{\boldsymbol{p}' \sim \mathbb{P}_g} \left[ D_\theta (\boldsymbol{p}') \right]
\label{eq:7}
\end{equation}
where $D$ belongs to the set of 1-Lipschitz functions, $\mathbb{P}_r$ and $\mathbb{P}_g$ are distribution of dataset and model distribution implicitly defined by $\boldsymbol{p}'=G_\phi(\boldsymbol{z}|\boldsymbol{C}) $, respectively. In addition, we employ the Wasserstein GAN gradient penalty (WGAN-GP) method to address the delicate and unstable training issues associated with vanishing or exploding gradients. Further details can be found in \citep{Gulrajani2017}.

\begin{figure}
    \centering
    \includegraphics[width=1\linewidth]{figures/3.pdf}        
    \caption{\textbf{Generative modeling framework}. The conditional WGAN model can address multiple conditions including property and geometry characterizations of TPMS unit cell, here we called it TPMS-WGAN. Each condition can be turned on$/$off to meet various design requirements. Considering elastic modulus, the TPMS-WGAN takes effective elastic modulus $E$ and random noise $\boldsymbol{z} (\boldsymbol{z}\sim\mathcal{N}(\mathbf{0},\mathbf{I})$,a Gaussian distribution of zero mean and the identity as the covariance matrix) as input to generator $G_\phi^1$, and the generated TPMS structure (denoted by implicit function parameters $\boldsymbol{p}'$), $E$ and the real TPMS structure (denoted by $\boldsymbol{p}$) are then passed to the discriminator $D_\theta$ to output value $f_D$. The $\boldsymbol{p}'$ also is passed navigator $R_\omega$ to obtained the predicted effective elastic modulus $E'$.}
    \label{fig:3}
\end{figure}

\subsection{Constraint-aware active learning}
\label{sec:2-4}
Active learning methods based BO are popular in addressing black-box problems that are expensive to evaluate. Unlike gradient-based and heuristic optimization techniques, which rely solely on the predicted values of surrogate models, BO utilizes the probabilistic predictions of a surrogate model that is updated sequentially. BO comprises two primary components: the surrogate model and the acquisition function (AF). Surrogate model make it possible to search the virtual design space at low computational costs with uncertainty, while AF is employed to evaluate the score of samples in a given virtual design space, considering exploration and exploitation. In the work, support vector machine (SVM) model is used to surrogates the expensive objective i.e., the FE calculation of yield stress of the TPMS unit cell. Then bootstrap method is utilized to introduce the prediction uncertainties to build the probability model $y=f(\boldsymbol{x})+\varepsilon$. The probability model is directly applied to estimate the mean value ($\mu$) and associated uncertainties ($\Sigma^2$) of candidates in a unexplored virtual design space.

BO is a sequential optimization process that explore the optimal solution through iteration. At each iteration, the SVM model is used to evaluate all the unobserved sample within virtual design space, and select the most promising candidate $\hat{\boldsymbol{x}}^*$ for simulation or experiment to obtain $f(\hat{\boldsymbol{x}}^*)$ according to AF. The obtained new data pair $(\hat{\boldsymbol{x}}^*, f(\hat{\boldsymbol{x}}^*))$ is added to the known data set $\mathcal{D} = \{(\boldsymbol{x}_1, f(\boldsymbol{x}_1)), \ldots, (\boldsymbol{x}_n, f(\boldsymbol{x}_n))\}$ and update SVM to enhance its prediction ability. The process is repeated iteratively until a predefined convergence criterion or budget is satisfied. The critical step is the selection of candidate point $\hat{\boldsymbol{x}}^*$ which is done via an AF that enables active learning of the objective $f(\cdot)$\citep{Settles2009ActiveLL}. AFs are denoted by the expected value of a user-defined utility function conditioned on the available data $\mathcal{D}$:
\begin{equation}
    \alpha(\boldsymbol{x}) = \mathbb{E}[I(\boldsymbol{x})|\mathcal{D}]
\label{eq:8}
\end{equation}
A popular utility function is expected improvement (EI) which rewards the large improvement over the best observed value $f^*$ in $\mathcal{D}$ (evaluated thus far), that is:
\begin{equation}
    I_\text{EI}(\boldsymbol{x})=\text{max}(f(\hat{\boldsymbol{x}})-f^*, 0)
    \label{eq:9}
\end{equation}
The corresponding AF can now by obtained by substituting Eq. \ref{eq:9} in Eq. \ref{eq:8} and using the reparametrization trick (more details are provided in Supplementaty S2):
\begin{equation}
\alpha_\text{EI}(\boldsymbol{x}) = \left(\mu(\boldsymbol{x}) - f^*\right) \Phi\left(\frac{\mu(\boldsymbol{x}) - f^*}{\Sigma(\boldsymbol{x})}\right) + \Sigma(\boldsymbol{x}) \phi\left(\frac{\mu(\boldsymbol{x}) - f^*}{\Sigma(\boldsymbol{x})}\right)
\label{eq:10}
\end{equation}
where $\phi(\cdot)$ is the probability density function (PDF), $\Phi(z)$ is the cumulative density function (CDF) of the standard normal random variable $z$.
The first term on the right-hand side addresses exploitation, while the second term is associated with exploration. The AF is employed to score the unknown samples in virtual design space, guiding the selection of potential candidates for next simulation or experiments.

Adding inequality constraints to BO can be formed by
\begin{equation}
\max_{c(\boldsymbol{x}) \leq \lambda} f(\boldsymbol{x})
\label{eq:11}
\end{equation}
where $c(\boldsymbol{x})$ are the other results of data set $\mathcal{D}$, and it is modeled by another probability model similar to $f(\cdot)$. The $c(\boldsymbol{x})\leq \lambda$ is a constraint function that limited the feasible area in virtual design space. The constraint-aware EI for a candidate $\hat{\boldsymbol{x}}$ can be written by:
\begin{equation}
I_{\text{EI-C}}(\hat{\boldsymbol{x}})=\Delta_\text{C}(\hat{\boldsymbol{x}})\text{max}\{0, f(\hat{\boldsymbol{x}})-f^*\}=\Delta_\text{C}(\hat{\boldsymbol{x}})I_{\text{EI}}(\hat{\boldsymbol{x}})
    \label{eq:12}
\end{equation}
where $\Delta_\text{C}(\hat{\boldsymbol{x}})\in \{0,1\}$ is a feasibility indicator function that is 1 if $c(\hat{\boldsymbol{x}})\leq \lambda$, and 0 otherwise. 
The distribution of random variable $\Delta(\hat{\boldsymbol{x}})$ is Bernoulli distribution:
\begin{equation}
    PF(\hat{\boldsymbol{x}}):=Pr[c(\boldsymbol{x})\leq \lambda] 
    \label{eq:13}
\end{equation}
The quantity $PF(\hat{\boldsymbol{x}})$ is a cumulative distribution function. These steps lead to the  constraint-aware expected improvement AF:
\begin{equation}
\begin{aligned}
\alpha_{\text{EI-C}}(\hat{\boldsymbol{x}}) &=\mathbb{E}[I_\text{EI-C}(\hat{\boldsymbol{x}})]\\
&= \mathbb{E}[\Delta_{C}(\hat{\boldsymbol{x}})I_{\text{EI}}(\hat{\boldsymbol{x}})|\hat{\boldsymbol{x}}]\\
&= \mathbb{E}[\Delta_{C}(\hat{\boldsymbol{x}})|\hat{\boldsymbol{x}}]\mathbb{E}[I_\text{EI}(\hat{\boldsymbol{x}})|\hat{\boldsymbol{x}}]\\
&=PF(\hat{\boldsymbol{x}})\alpha_{\text{EI}}(\hat{\boldsymbol{x}})
\end{aligned}
\label{eq:14}
\end{equation}
Thus $\alpha_{\text{EI-C}}(\hat{\boldsymbol{x}})$ is the a standard EI of $\hat{\boldsymbol{x}}$ over the feasible point weighted by the probability that $\hat{\boldsymbol{x}}$ is feasible. In the work, we extend the constraint-aware expected improvement AF by surrogate $PF(\hat{\boldsymbol{x}})$ with the normalized distance between the constraint value $c(\hat{\boldsymbol{x}})$ within a virtual space and a target point $c^*$. This distance is mathematically expressed as $ND(||c(\hat{\boldsymbol{x}})-c^*||)$. By incorporating the constrained function, the distribution of \(\alpha_{\text{EI}}(\hat{\boldsymbol{x}})\) is adjusted point-by-point. Notably, this approach eliminates the need to construct an additional probability model to surrogate the constraint function $c(\hat{\boldsymbol{x}}) $. Consequently, Equation \ref{eq:17} can be reformulated as $\alpha_{\text{EI-C}}(\hat{\boldsymbol{x}}) =ND(||c(\hat{\boldsymbol{x}})-c^*||)\alpha_{\text{EI}}(\hat{\boldsymbol{x}})
$.

Figure \ref{fig:6} schematically illustrates the comparison between the constraint-unaware active learning process and the constraint-aware active learning process. The optimal point in the constraint-unaware objective function is denoted by blue dot (Fig \ref{fig:6}a), while the optimal point in the constraint-aware objective function is denoted by red dot (Fig\ref{fig:6}b). We introduced a constraint function expressed by a normalized function to alters the EI score of candidates in a virtual data space, which is evident in the shift of the most promising point to query, as depicted from Fig. \ref{fig:6} c to Fig \ref{fig:6}d (more details about the toy problem are provided in Supplementary S3). The method can accurately found the optimal point in the constraint-aware objective function. Notably, the selection of the subsequent query point, whether driven by prediction (exploitation), uncertainty (exploration), or a trade-off between the two, will be influenced by the constraints.

@@

\subsection{Develop high yield strength TPMS unit cells with matched elastic modulus via the constraint-Aware Strategy}

>>>>

\title{Inverse design workflow of implicit surface-based cellular materials via conditional generative model and active learning}

\author[1]{Jiaxuan Ma}
\author[3]{Bin Cao}
\author[1]{Yuan Tian\corref{cor1}}
\ead{tian0321@shu.edu.cn}
\author[1,2]{Sheng Sun\corref{cor1}}
\ead{mgissh@t.shu.edu.cn}

@@

\begin{abstract}
Implicit surface-based cellular materials have gained significant attention in bone scaffold applications due to their structural resemblance to natural bone. These materials exhibit excellent mechanical properties and bioadaptability, providing an ideal environment for bone tissue regeneration. However, the rational design of such materials often relies heavily on expert knowledge and extensive trial-and-error experiments. In the work, we propose a data-efficient, two-stage bi-objective optimization framework within a machine learning paradigm, which accounts for the disparate computational costs associated with different objectives. Specifically, we apply this methodology to the inverse design of orthopedic implants, integrating a generative model, active learning, and finite element analysis to achieve cellular materials with both biocompatible elastic modulus and enhanced yield strength. Our approach demonstrates significant improvements in overall performance compared to the initial dataset, while reducing design costs by approximately sixfold relative to conventional one-step bi-objective machine learning optimization methods. This framework establishes an efficient paradigm for the rapid and intelligent design of cellular materials with tailored mechanical properties, particularly in scenarios where objectives involve imbalanced computational costs.
\end{abstract}

@@

%%Research highlights
\begin{highlights}
\item A TPMS-GAN with a navigator was developed to generate TPMS unit cells meeting specific mechanical and geometric conditions.
\item An inversed TPMS-GAN was introduced to design TPMS unit cells directly from geometric features in an explainable way, bypassing the need to interpolate in latent space or adjust implicit function coefficients.
\item A constraint-aware active learning method was proposed to tackle bi-objective optimization problems with imbalanced objective function costs.
\end{highlights}

@@

In the work, we propose a two-step strategy that integrates generative model with active learning algorithms to address bi-objective design problem of TPMS unit cells associated with imbalanced costs of properties. As demonstrated in Fig \ref{fig:1}, our approach consists of three main parts: (1) generate structures with a desired elastic modulus. In this step, a TPMS-GAN model is trained on a dataset containing TPMS unit cells and their corresponding elastic modulus. Then the generator of the TPMS-GAN produces structures with target elastic modulus, forming a virtual design space ($\Omega_E$). (2) inversed TPMS-GAN (ITPMS-GAN) model was developed to explore the latent space of TPMS-GAN, allowing for direct geometric manipulation of TPMS structures instead of adjusting implicit function parameters. (3) an elastic modulus constraint-aware active learning approach for yield strength optimization. Initially, the yield strength of TPMS structures selected from $\Omega_E$ are evaluated to establish the initial dataset via FEM. Subsequently, elastic modulus constraint-aware active learning based on Bayesian optimization (BO) is employed to optimize the yield strength of TPMS unit cells. BO method has been proven significant efficiency in materials design with limited data \citep{Ma2024b,Cao2024a,Tian2024a}.

\begin{figure}
    \centering
    \includegraphics[width=1\linewidth]{figures/1.pdf}
    \caption{\textbf{An overview of the proposed workflow for TPMS unit cells design.} \textbf{a} The generator in TPMS-GAN model produces virtual design space ($\Omega_E$) with desired elastic modulus. \textbf{b} The ITPMS-GAN model was developed to edit the TPMS structure through geometry features to rich diversity of candidates in $\Omega_E$. \textbf{c} The machine learning algorithm interactively queries the FE simulation to discover new TPMS designs with high yield strength and target elastic modulus through constrained active learning in $\Omega_E$.}
    \label{fig:1}
\end{figure}

@@

In the work, a voxel-based asymptotic homogenization method \citep{Dong2019} is employed to calculate the effective elastic tensor of TPMS unit cells. The asymptotic homogenization method posits that physical quantities such as stress and strain depend on both the macroscale ($X$) and the microscale ($x/\varepsilon$), where $\varepsilon$ is the scale factor. These quantities are homogenized at the macroscale and exhibit periodicity at the microscale. The microscopic strain tensor can be asymptotically expanded, and by neglecting higher-order terms, it can be expressed as:
\begin{equation}
    \varepsilon = \bar{\varepsilon} + \varepsilon^{*}
\label{eq:2}
\end{equation}
Here, $\bar{\varepsilon}$ represents the mean or macroscopic strain within the unit cell, while $\varepsilon^{*}$ denotes the perturbation strain at the periodic microscale. The perturbation strain is derived from the mechanical balance equation:
\begin{equation}
\int_{\Omega}\hat{\varepsilon}^T:\mathbb{C}:\varepsilon^{*}d\Omega = \int_{\Omega}\hat{\varepsilon}^T:\mathbb{C}:\bar{\varepsilon}
\label{eq:3}
\end{equation}
In this equation, $\hat{\varepsilon}$ is the virtual strain, $\mathbb{C}$ is the elastic tensor of the constituent materials, and $\Omega$ represents the unit cell. This equation is numerically discretized and evaluated using the FEM, which employs eight-node hexahedrons to voxelize each unit cell into $40 \times 40 \times 40$ cubes (see Fig S1). Once $\varepsilon^{*}$ is obtained, the effective elastic tensor $\mathbb{C}^H$ can be determined using the following equation:
\begin{equation}
\mathbb{C}^H:\bar{\varepsilon} = \frac{1}{|\Omega|}\int_{\Omega}\mathbb{C}:(\bar{\varepsilon}+\varepsilon^{*})d\Omega
    \label{eq:4}
\end{equation}
where $|\Omega|$ denotes the volume of the unit cell.

@@

In this matrix, $E_i$ represents the elastic modulus along the $i^{\text{th}}$ direction, where $i \in \{x, y, z\}$. In the work, for cubic symmetric unit cells, we have effective elastic modulus $E^H=E_x=E_y=E_z$. Additionally, elastic modulus in other directions can be obtained from $\mathbb{S}^H$ using the following equation:
\begin{equation}
\frac{1}{E_{ijk}} = \mathbb{S}^H_{11} - 2 \left(\mathbb{S}^H_{11} - \mathbb{S}^H_{12} - \frac{1}{2} \mathbb{S}^H_{44} \right) \times \left( \ell_{i1}^2 \ell_{j2}^2 + \ell_{j2}^2 \ell_{k3}^2 + \ell_{i1}^2 \ell_{k3}^2 \right)
    \label{eq:6}
\end{equation}
Here, $E_{ijk}$ is the elastic modulus along the $[i,j,k]$ direction, and $\ell$ is the cosine of the angle between the orientation $[i,j,k]$ and the coordinate axes $[x, y, z]$. This approach allows for the evaluation of elastic modulus in any orientation of the unit cell, facilitating the construction of a three-dimensional elastic modulus surface (see Fig S1).

@@

In the work, we employ a conditional Wasserstein GAN comprising four components: a generator network ($G_\phi$), a discriminator network ($D_\theta$),  a navigator network ($R_\omega$) and a condition module $\boldsymbol{C}$ that consists of mechanical properties and geometric features of the TPMS unit cell. Each condition can be turned on/off during the training process to meet various design requirements (see Fig. \ref{fig:2}). The generator $G_\phi$ (with set of trainable parameters $\phi$) map the Gaussian noise ($z$) and conditions to TPMS's parameter space $\boldsymbol{p}$ in $\mathbb{R}^4$, and achieve the fake $\boldsymbol{p}'$. While the discriminator $D_\theta$ (with set of trainable parameters $\theta$) evaluate whether a given set of structure parameters are "real" or not. It takes $\boldsymbol{p},\boldsymbol{p}'$ and  $\boldsymbol{C}$ as input and its output $D_\theta(\boldsymbol{p}'|\boldsymbol{C})$ is a function value representing the degree of realism of $\boldsymbol{p}'$. During the training process, the goal of the generator $G_\phi$ is to produce structure parameters $\boldsymbol{p}'$ of TPMS that can deceive the discriminator $D_\theta$. Simultaneously, the objective of $D_\theta$ is to distinguish structures $\boldsymbol{p}'$ generated by $G_\phi(\boldsymbol{z}|\boldsymbol{C})$ and real structure $\boldsymbol{p}$. $G_\phi$ and $D_\theta$ play the following two-player minimax game with loss function:
\begin{equation}
\mathcal{L}_{(G_\phi, D_\theta)}= \min_G \max_{D} \mathbb{E}_{\boldsymbol{p} \sim \mathbb{P}_r} \left[ D_\theta(\boldsymbol{p}|\boldsymbol{C}) \right] - \mathbb{E}_{\boldsymbol{p}' \sim \mathbb{P}_g} \left[ D_\theta (\boldsymbol{p}') \right]
\label{eq:7}
\end{equation}
where $D$ belongs to the set of 1-Lipschitz functions, $\mathbb{P}_r$ and $\mathbb{P}_g$ are distribution of dataset and model distribution implicitly defined by $\boldsymbol{p}'=G_\phi(\boldsymbol{z}|\boldsymbol{C}) $, respectively. In addition, we employ the Wasserstein GAN gradient penalty (WGAN-GP) method to address the delicate and unstable training issues associated with vanishing or exploding gradients \citep{Gulrajani2017}.

\begin{figure}
    \centering
    \includegraphics[width=1\linewidth]{figures/3.pdf}        
    \caption{\textbf{Generative modeling framework}. The conditional WGAN model can address multiple conditions including mechanical and geometric features of TPMS unit cell, here we called it TPMS-WGAN. Each condition can be turned on$/$off to meet various design requirements. Considering elastic modulus, the TPMS-WGAN takes effective elastic modulus $E$ and random noise $\boldsymbol{z} (\boldsymbol{z}\sim\mathcal{N}(\mathbf{0},\mathbf{I})$, a Gaussian distribution of zero mean and the identity as the covariance matrix) as input to generator $G_\phi^1$, and the generated TPMS structure (denoted by implicit function parameters $\boldsymbol{p}'$), $E$ and the real TPMS structure (denoted by $\boldsymbol{p}$) are then passed to the discriminator $D_\theta$ to output value $f_D$. The $\boldsymbol{p}'$ also is passed navigator $R_\omega$ to obtained the predicted effective elastic modulus $E'$.}
    \label{fig:2}
\end{figure}

\subsection{Constraint-aware active learning}
\label{sec:2-4}
Active learning methods based BO are popular in addressing black-box problems that are expensive to evaluate. Unlike gradient-based and heuristic optimization techniques, which rely solely on the predicted values of surrogate models, BO utilizes the probabilistic predictions of a surrogate model that is updated sequentially. BO comprises two primary components: the surrogate model and the acquisition function (AF). Surrogate model make it possible to search the virtual design space at low computational costs with uncertainty, while AF is employed to evaluate the score of samples in a given virtual design space, considering exploration and exploitation. In the work, support vector machine (SVM) model is used to surrogates the expensive objective i.e., the FE calculation of yield strength of the TPMS unit cell. Then bootstrap method is utilized to introduce the prediction uncertainties to build the probability model $y=f(\boldsymbol{x})+\varepsilon,\varepsilon\sim(0, \Sigma^2)$. The probability model is directly applied to estimate the mean value $f(\boldsymbol{x})$ and associated uncertainty $\Sigma^2$ of candidates in an unexplored virtual design space.

BO is a sequential optimization process that explore the optimal solution through iteration. At each iteration, the SVM model is used to evaluate all the unobserved samples within $\Omega_E$, and select the most promising candidate $\hat{\boldsymbol{x}}^*$ for simulation or experiment to obtain $f(\hat{\boldsymbol{x}}^*)$ according to AF. The obtained new data pair $(\hat{\boldsymbol{x}}^*, f(\hat{\boldsymbol{x}}^*))$ is added to the known data set $\mathcal{D} = \{(\boldsymbol{x}_1, f(\boldsymbol{x}_1)), \ldots, (\boldsymbol{x}_n, f(\boldsymbol{x}_n))\}$ and update SVM to enhance its prediction ability. The process is repeated iteratively until a predefined convergence criterion or budget is satisfied. The critical step is the selection of candidate point $\hat{\boldsymbol{x}}^*$ which is done via an AF that enables active learning of the objective $f(\cdot)$\citep{Settles2009ActiveLL}. AFs are denoted by the expected value of a user-defined utility function conditioned on the available data $\mathcal{D}$:
\begin{equation}
    \alpha(\boldsymbol{x}) = \mathbb{E}[I(\boldsymbol{x})|\mathcal{D}]
\label{eq:8}
\end{equation}
A popular utility function is expected improvement (EI) which rewards the improvement over the best observed value $f^*$ in $\mathcal{D}$ (evaluated thus far), that is:
\begin{equation}
    I_\text{EI}(\boldsymbol{x})=\text{max}(f(\hat{\boldsymbol{x}})-f^*, 0)
    \label{eq:9}
\end{equation}
The corresponding AF can now by obtained by substituting Eq. \ref{eq:9} in Eq. \ref{eq:8} and using the reparametrization trick (more details are provided in Supplementaty Note S2):
\begin{equation}
\alpha_\text{EI}(\boldsymbol{x}) = \left(\mu(\boldsymbol{x}) - f^*\right) \Phi\left(\frac{\mu(\boldsymbol{x}) - f^*}{\Sigma(\boldsymbol{x})}\right) + \Sigma(\boldsymbol{x}) \phi\left(\frac{\mu(\boldsymbol{x}) - f^*}{\Sigma(\boldsymbol{x})}\right)
\label{eq:10}
\end{equation}
where $\phi(\cdot)$ is the probability density function (PDF), $\Phi(z)$ is the cumulative density function (CDF) of the standard normal random variable $z$.
The first term on the right-hand side addresses exploitation, while the second term is associated with exploration. The AF is employed to score the unknown samples in virtual design space, guiding the selection of potential candidates for next simulation or experiments.

Adding inequality constraints to BO can be formed by
\begin{equation}
\max_{c(\boldsymbol{x}) \leq \lambda} f(\boldsymbol{x})
\label{eq:11}
\end{equation}
where $c(\boldsymbol{x})$ are the other results of data set $\mathcal{D}$, and it is modeled by another probability model similar to $f(\cdot)$. The $c(\boldsymbol{x})\leq \lambda$ is a constraint function that limited the feasible area in a virtual design space. The constraint-aware EI for a candidate $\hat{\boldsymbol{x}}$ can be written by:
\begin{equation}
I_{\text{EI-C}}(\hat{\boldsymbol{x}})=\Delta_\text{C}(\hat{\boldsymbol{x}})\text{max}\{0, f(\hat{\boldsymbol{x}})-f^*\}=\Delta_\text{C}(\hat{\boldsymbol{x}})I_{\text{EI}}(\hat{\boldsymbol{x}})
    \label{eq:12}
\end{equation}
where $\Delta_\text{C}(\hat{\boldsymbol{x}})\in \{0,1\}$ is a feasibility indicator function that is 1 if $c(\hat{\boldsymbol{x}})\leq \lambda$, and 0 otherwise. 
The distribution of random variable $\Delta_\text{C}(\hat{\boldsymbol{x}})$ is Bernoulli distribution:
\begin{equation}
    PF(\hat{\boldsymbol{x}}):=Pr[c(\boldsymbol{x})\leq \lambda] 
    \label{eq:13}
\end{equation}
The quantity $PF(\hat{\boldsymbol{x}})$ is a cumulative distribution function. These steps lead to the  constraint-aware expected improvement AF:
\begin{equation}
\begin{aligned}
\alpha_{\text{EI-C}}(\hat{\boldsymbol{x}}) &=\mathbb{E}[I_\text{EI-C}(\hat{\boldsymbol{x}})]\\
&= \mathbb{E}[\Delta_{C}(\hat{\boldsymbol{x}})I_{\text{EI}}(\hat{\boldsymbol{x}})|\hat{\boldsymbol{x}}]\\
&= \mathbb{E}[\Delta_{C}(\hat{\boldsymbol{x}})|\hat{\boldsymbol{x}}]\mathbb{E}[I_\text{EI}(\hat{\boldsymbol{x}})|\hat{\boldsymbol{x}}]\\
&=PF(\hat{\boldsymbol{x}})\alpha_{\text{EI}}(\hat{\boldsymbol{x}})
\end{aligned}
\label{eq:14}
\end{equation}
Thus $\alpha_{\text{EI-C}}(\hat{\boldsymbol{x}})$ is the a standard EI of $\hat{\boldsymbol{x}}$ over the feasible point weighted by the probability that $\hat{\boldsymbol{x}}$ is feasible. In the work, we extend the constraint-aware expected improvement AF by surrogate $PF(\hat{\boldsymbol{x}})$ with the normalized distance between the constraint value $c(\hat{\boldsymbol{x}})$ and a target point $c^*$. This distance is mathematically expressed as $ND(||c(\hat{\boldsymbol{x}})-c^*||)$. By incorporating the constraint function, the distribution of \(\alpha_{\text{EI}}(\hat{\boldsymbol{x}})\) is adjusted point-by-point. Notably, this approach eliminates the need to construct an additional probability model to surrogate the constraint function $c(\hat{\boldsymbol{x}}) $. Consequently, Equation. \ref{eq:17} can be reformulated as $\alpha_{\text{EI-C}}(\hat{\boldsymbol{x}}) =ND(||c(\hat{\boldsymbol{x}})-c^*||)\alpha_{\text{EI}}(\hat{\boldsymbol{x}})
$.

Figure \ref{fig:6} schematically illustrates the comparison between the constraint-unaware active learning process and the constraint-aware active learning process. The optimal point in the constraint-unaware objective function is denoted by blue dot (Fig. \ref{fig:6}\textbf{a}), while the optimal point in the constraint-aware objective function is denoted by red dot (Fig. \ref{fig:6}\textbf{b}). We introduced a constraint function expressed by a normalized function to alters the EI score of candidates in a virtual data space, which is evident in the shift of the most promising point to query, as depicted from Fig. \ref{fig:6}\textbf{c} to Fig. \ref{fig:6}\textbf{d} (more details are provided in Supplementary Note S3). The method can accurately found the optimal point in the constraint-aware objective function. Notably, the selection of the subsequent query point, whether driven by prediction (exploitation), uncertainty (exploration), or a trade-off between the two, will be influenced by the constraints.

@@

\subsection{Develop high yield strength TPMS unit cells with matched elastic modulus via the constraint-aware strategy}
